---
title: "Capital Delivery Model Data Analysis"
author:
- name: Arindom Baruah
  degrees: Master of Business Analytics, Monash University
  email: abar0090@student.monash.edu
  
email: BusEco-Econometrics@monash.edu
organization: Melbourne Water Corporation, Monash University
bibliography: references.bib
format: report-pdf
toc: true
header-includes:
  - \usepackage{tcolorbox}
  - \newtcolorbox{noteBox}{colback=yellow!5!white, colframe=yellow!50!black, title=Note}
  - \newtcolorbox{warningBox}{colback=red!5!white, colframe=red!50!black, title=Warning}
  - \newtcolorbox{tipBox}{colback=green!5!white, colframe=green!50!black, title=Tip}
  - \newtcolorbox{confBox}{colback=red!5!white, colframe=red!5!black, title=Confidentiality statement}
---


```{r}
#| echo: false
library(tidyverse)
library(kableExtra)
```
# List of abbreviations and terminologies

The analysis report will use numerous abbreviations and terminologies at various stages. This section will serve as a reference for what do these abbreviations mean alongwith their definitions.

1. __BNI__ : Business Need Identifier
2. __PBC__ : Preliminary Business Case
3. __BCA__ : Business Case Approval 
4. __Stage 1__ : First stage of project completion
5. __Stage 2__ : Second stage of project completion
6. __Stage 3__ : Third stage of project completion
7. __FFC__ : Final forecasted cost of project
8. __ID__ : Unique identification for each individual project
9. __Owner group__ : The division within Melbourne Water Corporation responsible for completing the particular project
10. __Delivery Program__ : The subdivision within Melbourne Water Corporation reponsible for completing the particular project


\newpage

# Executive Summary

The purpose of this study is to outline the methodology and guidelines for conducting the Capital Delivery Model project data analysis. This analysis aims to develop a reproducible framework that compares the current delivery model with past models, providing valuable insights to enhance future delivery models planned for years between 2026 to 2031.

By systematically examining past and present performance, the aim is to identify key trends, strengths, and areas for improvement, ensuring that future projects are executed more efficiently and effectively. 

For the purpose of the current analysis, Melbourne Water Corporation's project data with business need identifier (BNI) dates between 2008 to 2024 were analysed. The major areas of focus for the current analysis are delineated as follows:

1. Distribution of the projects in each delivery model based on the overall valuation of the projects as determined by the final forecast cost (FFC).
2. Distribution of the projects in numbers for each delivery model.
3. Distribution of project duration based on their overall valuations.
4. Project approval duration distribution across different stages of approvals.
5. Forecasted number of projects in the future delivery period from 2026 to 2031.

\begin{confBox}

To comply with Melbourne Water Corporation’s confidentiality requirements, the data has been de-identified and includes only the relevant fields necessary for obtaining insights for the intended evaluation purpose.

\end{confBox}

```{r}
#| caption: tbl-data
#| tbl-cap: "Project data"
project_data <- read_csv("data/Capital_Deliver_Project_deidentified.csv")

head(project_data) %>% kbl()
```


# Data Source

# Introduction

# Relevance of the current study

# Methodology

# Results and Discussion

# Conclusion

# Key areas of imrprovement